### 2024-03-19 writingdumpster.qmd
finished perceptual task section

---
title: "Writing Trash Bin"
author: "Harriet Mason"
format: html
---
# Why this exists
I have an information hoarding problem, this helps me throw out sections of the literature review that are not necessarily relevant without actually throwing it out. Knowing that I can come back to this at any point and pull information back, makes me more comfortable with removing it from the report.

# Citation Gaps
Cite Gap 1: When did uncertainty visualisaiton become its own subfield of visualisation?

Cite Gap 2: When was the first "uncertainty visualisation" used (likely a box plot of histogram)

Cite Gap 3: Some consider it to be synonymous with specific terms defined in mathematics, such as probability [*HERE*], variance[*HERE*], error[*HERE*], or precision[*HERE*]. Others consider it to be an encompassing umbrella term of which for these mathematically defined objects are only examples[*HERE*], and include concepts that are somewhat related, such as missing values[*HERE*]

Cite Gap 3.5: This wide array of uncertainty definitions becomes a snake eating its own tail, as many authors opt to redefine uncertainty themselves[*HERE*], ignore defining uncertainty entirely [*HERE*], or pick a definition from a seemingly randomly previously published papers [*HERE*], all of which make variance in uncertainty definitions even worse

Cite Gap 4:The uncertainty visualisation literature is completely awash with papers that define uncertainty using a vague encompassing phrase but go on to *quantify* uncertainty as a PDF, error, or some other easily quantified mathematical object when it comes time to do the visualisation.

Cite Gap 5: This definition swap happens so subtly that it seems to go unnoticed by authors writing the papers, and the fact that "uncertainty visualisation" methods dont even notice they are exclusively focus on depictions of easily quantifiable uncertainty, such as error [*HERE*], mass [*HERE*], or variance [*HERE*]

Cite Gap 6: This has left the expressions of uncertainty that are hard to quantify or visually differentiate untouched, despite many papers calling for their invention [*HERE*]

Cite Gap 7: Taxonomies split uncertainty based on an endless stream of arbitrary lines in the sand, such as whether the uncertainty is due to true randomness or a lack of knowledge [@Spiegelhalter2017 *HERE*], if the uncertainty is in the the attribute, spatial elements, or temporal element of the data [*HERE*], whether the uncertainty is scientific (e.g. error) or human (e.g. disagreement among parties) [*HERE*], if the uncertainty is random or systematic [*HERE*], statistical or bounded [], accuracy or precision [*HERE*], if the uncertainty is about the past or the future [*HERE*], the stage of the data analysis pipeline it comes from [*HERE*], the severity of the uncertainty [*HERE*], how quantifiable the uncertainty is [@Spiegelhalter2017], and many, many, many more examples.

Cite Gap 8: It is commonly noted that each uncertainty must be discussed in isolation, but combining the uncertainty from every stage is near impossible [@Spiegelhalter2017 *MORE HERE*]

Cite Gap 9: Many papers will boast a title that claims to be about uncertainty visualisaiton, but simply depicts different visual representations of a PDF, however it takes reading the methodology to find this out

Cite Gap 9: For example @Leland2005 mentions that popular graphics, such as pie charts and bar charts omit uncertainty, however at least one or both of these charts are used in most "uncertainty visualisation" experiments [@Ibrekk1987 *MORE HERE*]

Cite Gap 10: Uncertainty Visualisation Taxonomies

Cite Gap 11: Papers will often discuss uncertainty in relation to one of these motivating reasons, the evaluation experiments motivated by (1) often perform different visualisations of PDFs, while the papers motivated by (2) will focus on trying to impute uncertainty as error within one of the existing channels, such as colour

Cite Gap 12: Visualisation authors are almost unanimous in commenting that the "information" in two plots must be the same in order for the visual techniques to be compared

Cite Gap 13: Examples of what is "information" in a visualisation

Cite Gap 14: Examples of mathematically obvious results

Cite Gap 15: Examples of replicated perceptual task experiments

## Extra Section Cite Gaps
Cite Gap - New Section: Logical fallacy interpretation of mathematically unanswerable questions. If I tell you my cousins age, and then ask you to guess my age, it would not be unreasonable to guess the age I provided for my cousin. Citing this work to state "people incorrectly believe my cousins age is my age" would
- Deterministic questions about a random event cannot be answered. 
- @uncertchap2022 took the results of other paper to mean "when participants viewed mean temperature forecasts that included 95% confidence intervals depicted as bars with end caps, they incorrectly believed that the error bars represented high and low temperatures" (this one will depend on that study)

# Moved over notes
- @Blenkinsop2000 asked different questions for each uncertainty visualisation, indicating they had some awareness that the question has a significant impact on whether or not the visualisation should be used.
- @Gschwandtnei2016 When answering “What is the statistical probability (in %) that the interval has already ended at the marked point in time (red line)?” many answers gave the estimated probability to which the marked point in time falls into the interval. This is correct when asking for the probability to which the interval has already started at the marked position. However, in case of the probability to which the interval has already ended at the marked position, this answer is the inverted probability of what we were looking for. They believe this is their own mistake so they so they don’t present these results.

@Sanyal2009 users performed significantly well using Glyph-size when the task was to search for locations of least uncertainty. However, both Glyph-color and Surface-color performed better than Glyph- size when the task was to search for locations of high uncertainty. We did not expect to find a significant difference between the two search tasks since both the tasks were designed to find extremes in the data
@Kim2019 believed an understanding of participants prior beliefs could assist in interpreting uncertainty visualisation results that tend to be incredibly noisy.
@Thomson2005 titled their paper "A typology for visualizing uncertainty" however it is a taxonomy for the specific types of uncertainty that we may want to visualise and a way to calculate that uncertainty. While this paper is more about mathematics than visualisation, it is clearly written as a precursor to visualisation.
- Some markers of probability that have common uses, probabilities such as 0 (impossible), 0.5 (a fair coin flip), 1 (certain) are easy for people to have an internal sense of *hullman paper*. 
- It is incredibly hard to communicate small risks, because people cannot differentiate between 1/100 an 1/1000, and communicating low probability but high impact risks are particularly challenging [@Spiegelhalter2017]. This issue can be somewhat alleviated by communicating the likelihood of an event in relation to another event on a similar scale (such as being struck by lightning) [@Spiegelhalter2017]. 
- The framing of how we communicate risks matters due to the affect heuristic, so there will be a difference in peoples behaviours if, for example, you tell people they have a 90% chance of survival vs a 10% chance of death, and when expressing a frequency, a larger numerator communicates a larger risk [@Spiegelhalter2017]. 
- The reference class of a particular risk should be explicitly stated to avoid confusion, for example if we are communicating the chance of rain, the time period over which we are expecting this chance of rain (i.e. 50% chance today or this week) needs to be specified [@Spiegelhalter2017]. 
- @Munzner2009, in her nested model for visualisation, comments on how even something as low level as poor problem abstraction can cause down stream effects and become impossible to ignore. 
- @Otsuka2023 highlights that the choices made in the ontological process of boiling real world events down into statistical objects can influence later stages of analysis due to what was considred signal and what was considered noise. 
- @Meng2021 stresses the importance of treating data as an input from another process with noise and variance, rather than as a neutral input.
- @Griethe2006 clearly sees uncertainty as additional information or data, and therefore sees the uncertainty visualisation problem as a specific case of the high dimensional visualisation issues. 
- @Griethe2006 mentions that there are detailled classifications of uncertainty but the literature and subsequent techniques are not ballanced, and instead concentrate on scalar values. In facing this lack of ballance she creates a new classification system for uncertainty visualisation. 
- @Griethe2006 commented that uncertainty appears at every step in the analyis pipeline, and commented on some the mathematical definitions that seek to quantify uncertianty, but did not explain how these mathematical definitions can be combined across levels. Unfortunately this quantification is needed since she mentioned that in order to visualise uncertainty, it needs to be measurable.
- @Griethe2006 defined uncertainty as "the degree to which the lack of knowledge about the amount of error is responsible for hesitancy in accepting results and observations without caution". 
- @Griethe2006 found that there there are two ways to visualise uncertianty, implicitly and explicitly. When a visualisaion is implicit, the uncertainty is considered an uninteresting feature of the data, and instead of mapping some quantification of uncertainty to a visual feature, data is filtered out if it does not meet a specified "certainty" threshold. These statements clearly points towards a signal/noise dichotomy with normal visualisation and uncertainty visualisation.
- @Spiegelhalter2017 even if most audience members can read a plot, they may not read it in the same way. People with high numeracy skills count pictographs, but people with low numeracy assess area, and seemingly unrelated elements may not equally influence all these methods of reading, for example scattering in a pictograph makes counting particularly more difficult. *also add that perceptual task audience thing*
- @Spiegelhalter2017 it is better to understand your audience and iterate towards a final design, rather than only consider one option.
- @Spiegelhalter2017 common sentiments such as avoiding chart junk, considering a table in place of a visualsiaiton, use of narrative labels, managing emotion, less is more approach, etc, are the main concepts expressed. - @Spiegelhalter2017 suggests that the graphic should allow "part-to-whole" comparison on an appropriate scale, which seems to be an extension of the framing concept discussed when speaking about uncertainty communication more broadly.
- @Hullman2016 commented on the lack of justification of particular questions used in experiments and the lack of generalisability in using plot ranking as a response technique.
- @Hullman2016 discussed the lack of clarity in exactly what "uncertainty" authors are trying to ask about. Whether the participants are supposed to answer according to some ground truth or their subjective probability is unclear, as is whether the *uncertainty* depicted represents the uncertainty of a sampling or population distribution. 
- @Hullman2016 discussed the role of heuristics in the analysis of results and suggested authors look for signs of them in responses and ask subjects to describe their strategy in their response. 
- @Hullman2016 What was particularly interesting was people's tendancy to ignore uncertainty in a plot all together, and use a pattern in the signal (such as distance between means) to answer questions that require uncertainty information. 
- @Kinkeldey2014 stress the importance of selecting the correct audience, as numerical and visual literacy can differ vaslty between subpopulations so the results from one populations may not reflect the results from another. 
- @Daradkeh2015 found that people tended to to make decisions on a small number of salient pieces of information rather than on systematic evaluations of the costs and benefits of choices.
- @Daradkeh2015 found that people tended to to make decisions on a small number of salient pieces of information rather than on systematic evaluations of the costs and benefits of choices.
@Daradkeh2015 found that people tended to ignore or give little weight probability information and instead focused on the severity of undersierable outcomes.
@Daradkeh2015 found that people tended to ignore or give little weight probability information and instead focused on the severity of undersierable outcomes.
- @wu2023rational attempted to remedy this issue caused by designing a "rational agent benchmark" such that participants in an experiment are compared to a rational agent. This suggestion is an insufficient band-aid solution that is built upon the same invalid assumptions as the rest of the literature, and simply digs the field deeper into these problems, rather than solve them. 
- @wu2023rational almost seems to recognise the problem of decision making experiments with payment schemes or correct/incorrect answers, but instead of understanding this boils these tasks down to value extraction, they suggest participants should be compared to a "rational agent". This rational agent acts as a benchmark, in the same way a ground truth statistic does for value extraction experiments, except the authors don't seem to realise that is what they have done. This paper highlights many decision making tasks are simply a value extraction experiment without realising it, this paper included.
- make sure quantified vs discussed uncertainty is mentioned
- industrial accidents have a "tollerable" risk chart based on risk and number of fatalities @Spiegelhalter2017
- @Suh2023 found that the main issue with deployment of AI was communication between data science and subject matter experts, since subject matter experts didn't want technical presentations that didn't explain exactly when and how a model would fail.
- People believe in the law of small numbers but dont believe in the lar of large numbers @Hullman2016
- experiencing an event makes your percieved probability higher than if you see a description of it @Hullman2016
- Frequency thing means that more effective communication to participants in a study means more effective science and we have a roll on effect @Hullman2016
- Probability judgements are less relative than other psychophysical judgements, 0%, 50% and 100% all have well understood meanings. Subjective probability distributions are spontaneous, and subjects are constructing a distribution upon being asked rather than articulating a fully formed mental representation of their beliefs. Therefore it is subject to heuristics. consider anchoring a scale with farmiliar probabilities . not sure if being asked for the entire distribution helps, e.g. the graphical one reduces noise but asking people for an interval estimation leads to overprecision and overconfidence @Hullman2016
- believed that offering enough decisions with varying rewards you can observe the true subjective probability, however whether or not probability is the right representation of subjective uncertainty is a topic of debate @Hullman2016
- she also discusses graphical interface (users have to draw the probability distribution) and framing probabilities as natural frequencies are better choices because they reduce noise and stuff.@Hullman2016
- Consider incorperating utility functions. Include "probability-coherence" checks @Hullman2016
- evidence that people do not follow the axioms of probability when conveying subjective probabilities. How do you deal with this as an analyst (e.g. if the probabilities do not sum to 1 @Hullman2016
- we consider a summary table as a visualisation, illuminate graphics with words and numbers, design graphics to allow a part-to-whole comparison on an appropriate scale, use images and metaphors to gain/retain attention but dont arouse undue emotion, assume low numeracy of a general public audience and adopt a less-is-more approach by reducing the need for inference and making clear and explicit comparison, be cautious about interactivity and animations (may introduce unnecessary complexity), avoid chart junk (like 3d bar charts) @Spiegelhalter2017
- @Maceachren2012 found to be the worst aesthetics to map to uncertainty to as they don't have an intuitive interpretation.
- @Bella2005 found that most participants were ignorant to the fact that error bars are used for both confidence intervals and standard error bars, two wildly different indicators of precision. (**also the paper sherry sent me**)
- Several studies probide evidence that the usability of uncertainty representations can be highly user and task dependent @Kinkeldey2014
- The tasks for participants were most commonly value retrival (some where uncertainty and data value are retrived separately, others where they are some separately). also also be asked to extract a relative value @Kinkeldey2014
- most studies used coincident approaches, static visualisations, are applied to a specific domain and dont have generalised results @Kinkeldey2014
- Most studies involve intrinsic approaches such as colour, transparency, and those that used extrinsic techniques typically used glyphs, error bars, grid based techniques, or contouring @Kinkeldey2014
- Mentioned alleatory vs epistemic uncertainty, in the visualisation literature alleatory uncertainty is more commonly the focus. also mentions dependence and joint pdfs as a consideration @Hullman2016
- When asking questions, researchers should focus on events that can be repeated (because of a misunderstanding of confidence intervals) @Hullman2016 (moved here because it is clear she is touching on the whole "what are we uncertain about" thing)
- we should aim to show enough information to solve a task while avoiding irrelevant distracting information [@kosslyn2006graph]. 
- people will discount information they perceive as discountable if they are able. Including mean estimates on depictions of mass can cause people to discount the uncertainty information and use the difference between means as a proxy for the probability distribution [@Kale2021].
- rarely used extension is aggregation of uncertainty over an area (retrieve overall estimation from a spatial distribution of uncertainty). Other tasks include comparisons, rankings. also have a task called "search" where participants have to identify entities that fulfill certainty characteristics (high or low values) @Kinkeldey2014
- Whether to use classed or unclassed uncertainty is rarely discussed @Kinkeldey2014. e.g. land map with remotely sensed images has ambiguity in the boundaries, which can result from multiple sources (land cover class, measurement error, images from different dates). If you combine them all it might not make sense for someone who only needs one.
- The way we ask questions and the types of questions we ask are selected with little justification. This paper makes suggestions to reduce the noise in the data from these papers. @Hullman2016
- Error is best mapped to fuzziness, location, and colour value; arrangement, size and transparency are an OK second choice; but saturation, hue, orientation and shape are unacceptable and have no intuitive connection to variance [@Maceachren2012]. 
- @Nathonours found that a scatter plot is better than a line plot if you want to convey the correlation between two time series but we cannot be sure if this was influenced by swapping the distribution depicted or by dropping the irrelevant feature of inexchangeability.
- No only do the graphical elements we map our features to matter, but the direction matters too. Graphical elements that are more fuzzy (fuzziness), further from centre (location), lighter (colour value), poorly arranged (arrangement), smaller (size), more transparent (transparency) are perceived to be more uncertain [@Maceachren2012]. 
- This idea also extends to interval estimation, where questions about probability are best answered with gradients and questions about start and end times are easiest to answer with ambiguation [@Gschwandtnei2016].

- mentions uncertainty can be variance, precision, accuracy, reliability, or related concepts @Hullman2016
-  @Spiegelhalter2017 defines risk as risk known probabilities and uncertainty as estimated probabilities, therefore uncertainty is what we have when you cannot accurately quantify risk.
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016 
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016
- The big data paradox shrinks confidence intervals but magnifies bias [@Bradley2021] because the confidence intervals shrink due to the sample size but the imprecise target group creates bias.
- coloured maps for earthquake risk are easily interpretable @Spiegelhalter2017
- Hurricane risk is not only the path, but also the storm surge and wind speed. @Spiegelhalter2017

# Experiment paper checklist
General
- Specific Field of Research
- Question asked
- Risk or Uncertainty: 
- Noise vs signal (risk vs uncertainty)
- Major task goal
    - e.g. identify a bias or misunderstanding, compare two visualisation methods
- Minor task goal
    - Communication(value retrieval)
    - Analytical(complex consideration)
    - Exploration(free for all)

Uncertainty Considerations
	- Source
	- Level (statistical-ignorance)
	- Nature (Epistemic/Aleatory)

Visual Uncertainty taxonomy
- Explicit/implicit
    - (directly mapping or showing multiple outcomes)
- Intrinsic/Extrinsic 
    - (using existing symbols e.g. colour value, or new objects e.g. grids)
- Visually integral/separable 
    - (can or cannot be separated from the data) 
- Coincidence/adjacent 
    - (if data and uncertainty are in integrated or separate views
- Static/dynamic 
    - (animation/interaction)

Other plot considerations
	- Dimension of data
	- Dimension of uncertainty
	- Feature mappings

Extra (not necessarily recorded)
	- Metrics used (and recorded)
	- Possible Heuristics
	- What is the ground truth?
	- Participant literacy

# Main section outtakes
These results provide some insights that applies to visualisation more broadly but nothing specific to uncertainty informaiton. It is clear that audiences primarily interact with visualisations through their "System 1" brain [@daniel2017thinking]. This aligns with other authors comments that visualisation is primarily about "gists" [Spiegelhalter2017]. In this sense, asking participants a question that requires them to shift to their "System 2" is entirely contrary to *why* we use visualisation. The exact reasons we describe visualisation to be so powerful, that they are able to communicate complicated information fast and efficiently *is a product of the system 1 brain* [@daniel2017thinking], however this means displaying information using visualisations have the same weakness' as system 1 thinking, and visualisation authors need to be aware of this. Asking participants to answer questions that require shifting to their system 2 brain is not in the spirit of visualisation, and does not reflect how visualisations will be used in practice, and will very likely result in incorrectly answered questions. 
There is a large number of examples where participants got "confused" but the problem is more likely that the question required participants to shift to their system 2 brain.

There is also the common view that uncertainty as a dimensionality problem. This perspective is not quite right and the literature seems to be about 50/50 on whether uncertainty is a psychological or dimensionality problem.  It is important to keep in mind that just because information is in a graphic, that does not mean it will be "seen". The phenomena of inattentional blindness shows that there is no perception without attention and it is powerful enough that participants can fail to be aware of random objects appearing on a screen or a gorilla walking through a basketball game [@simons1999gorillas; @mack2003inattentional]. Therefore if information is mapped to graphical elements that are so low on hierarchy they can be ignored, they might as well not be there at all. The people who don't plot uncertainty because they think its unimportant and the people who relegate uncertainty to the lowest ranking aesthetic on the list are both saying the same thing, "I think the uncertainty is unimportant". Including uncertainty is worth very little if no attention is left to see it. This does not mean we cannot put a *large* amount of information in a graphic. Glyph maps can be used to depict the multi-dimensional information in spatial-temporal data by mapping line plots to map locations, but we still need to decide what information is important. The map can either present trends in the global or local variance depending on whether or not the line plot is scaled globally or locally [@Wickham2012], however smaller details are almost impossible to convey. No matter how much information we try and put in a plot, there will always be only a handful of key takeaways.

This issue has come up in many different forms in the uncertainty visualisation literature. Typically as a "task dependence" on the performance of uncertainty visualisations. It is repeatedly identified as a problem in previous reviews of the uncertainty visualisation literature [@Kinkeldey2014; @Hullman2016] as well as across many sub domains and applications [@Wallsten1997; @Fischhoff2014; @Meng2021; @Amar2005]. The fact that this conclusion is repeatedly reached shows both the importance and the lack of acknowledgement this concept receives. @Fischhoff2014  takes this approach by considerting how decisions with different levels of flexibility should have different uncertainty visualisations. If the user has no choice they should simply be made aware of the risk, if they have several options, the user needs to see the uncertainty in each, if they need to come up with the options themselves, they need to see how things work and how the system fails. @Kinkeldey2014 suggested that the nature of the task plays an important role and believed the task dependence for uncertainty visualisation could explain some of the inconsistency in outcomes.

Several uncertainty visualisation literature reviews have commented on the immense difficulty of combining results across experiments, but none have been able to pinpoint the reason. @Kinkeldey2014 suggested a task based taxonomy that organised the plots by the question or purpose of the visualisation, and even suggested that communication tasks (such as value retrival of uncertainty values) should just follow traditional cartography rules, @Spiegelhalter2017 simply ignored the existence of uncertainty visualisation as a subfield and just suggested we follow established rules within uncertainty visualisation, @Hullman2016 said that commented on the large amount of noise generated from uncertainty visualisation evaluation methods that led to a difficulty in combining results, @Griethe2006 did not meaningfully combine the work and instead listed findings in the field, but also did not comment on the lack of generalisable results. While we largely agree with the difficulties of these authors, we do not agree that the solution is a task based taxonomy or better questions. While we agree the lack of attention to tasks and appropriate questions has generated problems in the field, there is a much more pressing issue, which is that uncertainty visualisation authors don't seem to understand *why* uncertainty visualisation different to typical visualisations.

Attempts to rectify these conflicts always flutter about the surface and do not seem to grasp the core of this problem.  What these suggestions fail to understand is that the entire practice of identifying "true" facts about "unertainty visualisation" is futile. It is impossible to make a statement about the properties of an "uncertainty visualisation" in the same way it is impossible to comment on $\frac 0 0 = 1$. Division by zero is undefined, so the statement is ill formed, similarly, "uncertainty" is not defined, so statements *about* "uncertainty" including statements about "uncertainty visualisation" are ill formed. Ill-formed statements can neither be true nor false and cannot be part of a logical argument, so experiments and research that build upon this ill-formed concept are neither true nor false, they simply exist. Someone who is less familiar with the rules of logical argument may believe this understanding of the problem to an insufferable nitpick that creates impossible to live up to standards, however the consequences of this definitional issue are *very* real and costly.  The studies that investigate "uncertainty visualisation" may be identifying *something*, however there is good reason to believe that *something* is neither interesting nor does can it contribute to a larger theory of visualisation. This is something we illustrate in the following this section.

- The process of working out whether or not a visualisation shows a piece of information, and if it does, what perceptual task it is mapped to and what assumptions or heuristics need to be used, is *incredibly* time consuming. While this method does predict the results for these experiments with remarkable accuracy, to do it for a large number of plots, especially when the ground truth statistics and visualisation are incredibly complicated is just not feasible for a reviewer to do.
- This is likely because The idea that uncertainty can only be defined in the presence of a motivating question is well grounded in most areas of statistics. The entire process of data analysis, from deciding what should be observed as data through to communicating that data in a plot is governed by human decision and the goal of an analysis.
- We know uncertainty cannot be entirely explained by individual concepts that *have* been defined, such as probability, variance, error, missing values, precision, or related concepts that pop up throughout the data analysis pipeline. We also know it likely has some relationship with the *task* we are considering, given that many authors have attributed the noise in uncertainty visualisation research as a type of "task" dependence. Additionally, it should be clear that we wouldn't consider uncertainty to be an appropriate term for *everything* in statistics would we?
- What exactly is uncertainty, then? Let us consider the concepts related to uncertainty that have been discussed so far. We know uncertainty cannot be entirely explained by individual concepts that *have* been defined, such as probability, variance, error, missing values, precision, or related concepts that pop up throughout the data analysis pipeline. We also know it likely has some relationship with the *task* we are considering, given that many authors have attributed the noise in uncertainty visualisation research as a type of "task" dependence. Additionally, it should be clear that we wouldn't consider uncertainty to be an appropriate term for *everything* in statistics. For example, I wouldn't describe the population parameters of a distribution I have simulated to be "uncertain"

- tasks e.g. "What is the average of X" (how well lined up you are moving down the data pipeline) and resolution (uncertainty is very often moving back a resolution level in statisitcs - also can be an assumption - bootstrap is a replacement for an assumption, not a resolution change, but bootstrapping is centered on the data)  e.g. "What is the noise around the estimate of the value of X"
Visualisation experiments are highly sensitive to the way questions are asked. For example, a study by @Hullman2015 that compared HOPs, error bar, and violin plots found that many participants reported that group A was more likely to be higher than group B, which was not not plausible given that the mean of B was larger than the mean of A. The authors attributed this obvious mistake to the visual representation, but we believe it was more likely caused by the distributions being shown in the order A then B while the question swapped the order by asking "What is P(B>A)?". This flip is minor, but must be done *after* the participants have worked out the probability. Since the primary task is done they forget to complete the secondary task, similar to forgetting your bank card in an ATM. I suspect this was the cause of the mistake as I, when attempting this question myself, made the exact same mistake. I also constalty lead my bank card in ATMs.

While mathematical information *should* be considered the starting point in considering if two visualisations are equivalent, it is not the only consideration. A large amount of work in perceptual tasks, attention, and psychology surrounding charts shows that all the information depicted is not paid attention to in equal measure. Elements such as colour can have sizable impacts on the way a graphic is percieved. Mathematics identifies the information that is *shown* while psychology identifies the difference in information *recieved*. A failure to understand this distinction leads to graphics that are different not only in the way information is depicted but in the information itself. This problem is pervasive in uncertainty visualisation literature.

This is not to say any paper that discusses uncertainty visualisation is completely pointless because the results already exist. We are simply pointing out that treating uncertainty graphics as special simply because they contain uncertainty, even if that uncertainty is not related to the task, leads to results that are already established in information visualisation.

These two perceptualisations of *why* uncertainty visualisation is different from normal visualisation is also what leads to differing opinions on what is an uncertainty visualisation. For example, someone who percieves of uncertainty visualisation as a dimensionality problem would see a pie chart as a normal graphic, and a pie chart with blurred boundaries as an uncertainty visualisation (even though blurring the boundary between segments of a pie chart makes no sense when you think of a pie chart as a polar co-ordinate bar chart), while someone who sees uncertainty visualisation as psychologically different would see a pie chart as an uncertainty graphic so long as it depicts probabilities. This may lead you to believe that those who believe uncertainty is unique due to its psychological properties believe uncertainty should be defined at the mathematical stage, while those that believe it is an addition to an established visualisation think uncertainty is defined in a graphic at the aesthetic stage. This is not the case. As shown by the *grammar of graphics fig* there is no single person or experiment that is consistent in any one belief. This difficulty and lack of consistency comes from the desire to discuss "uncertainty visualisation" as though it is a single term and not the combination of "uncertainty" and "visualisation". If we split the terms and look at the quantification and visualisation of uncertainty separately, the results of many uncertainty visualisation papers become self evident.

Papers that are justified using the first type of motivation rarely explain exactly how the psychological apsects of uncertainty impct the participants in a way that inhibits their ability to read information from a graphic. Since uncertainty is poorly defined and the psychological effects of uncertainty is never justifiably connected to the scientific communication of it, as was discussed in ssections 3 and 4, it is unclear exactly *what* these papers are testing for. It is almost as though they *feel* that uncertainty visualisations *should* be different, but do not consider *how* uncertainty information is supposedly different beyond decision making criteria such as *risk avoidance* which is not something science communicators have any business trying to control, as was discussed in previous sections. For these reasons, experiments that test for secondary factors (such as influence on decision making) are confounding far too many variables to establish any generalisiable take away, and experiments that ask participants to extract values fail to notice there is no longer a reason for "uncertainty" to be any different to any other type of visualisation.

Papers that are justified using the second type of motivation clearly fit within the sub-field of high dimensional visualisation, but also rarely comment on the connection. These experiments evaluate different ways to include even *more* information than would be put in a typical visualisation. Some of these methods, such as a bivariate pallet or glyph map, create graphics that are overloaded with information and offer incorrect or meaningless takeaways at first glance. This is not to say *every* visualisation method that sees it as a dimensionality problem is bad. Clever methods of combining several types of information, such as value supressing uncertainty pallets (VSUPs), can better help us understand how our brains process information in a graphic.  For example, the VSUP allows us to test is colour value and saturation can be combined such that we mentally combine signal and noise together as one variable. Since the take aways from the papers are not usually related to uncertainty specifically, and often related to interesting perceptual tasks, there is also no justification for why this work is "uncertainty visualisation" work and not simply perceptual task studies that can be used by the rest of the information visualisation community.

This stems from a class of "uncertainty visualisation" that seems to have no rules for what belongs in the group, and no consideration for the mathematics that is actually depicted in these plots. Uncertainty visualisation experiments do not work to build upon existing visualisation literature or differentiate uncertainty visualisaiton as a new field, but seem to exist with shocking indifference to the field it belongs to.

uncertainty visualisation, as it is now, should not exist. Any generalisable contribution made by these papers is already an established result in information visualisaiton, and therefore is not specific to "uncertainty visualisation".  Any "uncertainty visualisation" specific result is not generalisable beyond the experimental conditions because of the poor formalisation in the definitions of uncertainty and uncertainty visualisations.

This lack of formalisation is not unique to uncertainty visualisation and the wider field of information visualisation suffers form similar problems [@Kinkeldey2014].

This makes it clear we have a serious problem with the classification of "uncertainty visualisation". The visualisations follow identical rules to those established already in the visualisation literature, and experiments that view uncertianty visualisation as somehow special or different from normal visualisation are simply repeating existing established research in information visualisation.

If several graphics are compared that are different statistically, geometrically, and aesthetically, there is *no way to combine the results to establish generic principles for the field*.

The overwhelming majority of these papers depict different mathematical expressions of both the signal and the uncertainty. In 100% of these papers, the plot that has mapped the statistic in question to some visual feature will outperform those that don't. When the statistics used to depict uncertainty *are* the same (which is rare) 100% of those papers will follow the hierarchy of perceptual tasks that has already been established in the visualisation literature. If the experiment participant needs to extract a probability and one graphic uses position along a common scale, while the other uses colour or literally any other perceptual task, the position depiction will always outperform the other.

If we were to compare two different visualisations, we would like to change as few features as possible so we can better understand the cause of the difference in perception.

Additionally, if information that can be used to calculate a statistic, but not the statistic itself, is shown, people will need to use a heuristic to extract that information which will decrease the precision with which that information is communicated. If the information that is depicted on a plot cannot be used to *mathematically* generate a best estimate for a specific statistic, it is questionable how uncertainty visualisation authors expect participants to cover that gap if *not* for heuristics.

In these papers will typically ask separate questions about the signal (e.g. what is the mean of this distribution) and the uncertainty (e.g. what is the variance of this distribution).

Failling to understand this key aspect of visualsiation leads to graphics that *contain* the correct amount of information, but that volume of information cannot be readily accessed by the viewer.

## Taxonomies
Some taxonomies contain overlapping terminology but also have alternative definitions and categories. For example *___* organises uncertainty into the two categories, Aleatroy (inherrent randomness) and epistemic (lack of knowledge), but *___* organises uncertainty into the three categories aleatory (inherrent randomness in a forecast) vs epistemic (uncertainty in structure and parameters of statistical models) vs ontological (uncertainty about the entire modelling process as a description of reality).

To avoid alienating lay people with confusing complexity, many of these taxonomies are overly simplistic. Most taxonomies seek to organise uncertainty based off where the uncertainty is coming from. For example, @Spiegelhalter2017 organises uncertainty into three categories: 
  - Aleatory uncertainty: inevitable unpredictability due to unforeseeable factors 
  - Epistemic uncertainty: uncertainty about the structure and parameters of statistical models
  - Ontological uncertainty: uncertainty about the entire modelling process as a description of reality.
These three categories can be considered to be uncertainty when deciding which model we should use (ontological), uncertainty surrounding parameter estimates (epistemic), and uncertainty surrounding a forecast (aleatory), however it is not hard to find another, very different, example of a taxonomy that organises uncertainty based on what is seen to be causing it. @Gustafson2019 organises uncertainty into the four groups:
  - Deficient: Lack of knowledge
  – Technical: modelling approximations and measurement error
  – Scientific Unknown unknowns
  – Consensus: Disagreement among parties
However the additional comments about the inherent variability of aleatory uncertainty results in a bit of confusion as to if our cateogirisation should change if we decide to treat the parameters of a statistical model to be inherently random (as is done in Bayesian statistics) or if we beleive a forecast variability to be caused by a lack of knowledge.

He kept the taxonomy to three levels to maintain simplicity, however this organisation methods conflates amount of uncertainty present in an estimate with what is causing that uncertainty, and in doing so makes it hard for us to use this definition to better understand uncertainty within our own analysis.
# context section outtakes
Given that uncertainty estimation and communication seem to be largely task depenent, what questions should we be testing? This leads to several unanswered questions. Is our goal in uncertainty communication to acknowledge and convey the inherrent randomness that exists in all statistical calculations and models? Is it to supress signal that could be the result of random variance? Is it to find the best way to convey the mass function of a random variable? A survey of the literature will leave the answers to these questions unclear, however this inherrent issue has not gone unnoticed.
This conceptual way of thinking about uncertainty has been rattling around in the different areas of data science for a while.
For example, when we record coin flips we typically want to model the behavior with a binomial distribution, so we ignore the outcome of a coin landing on its side despite the fact it is a real world possibility.

# General Thoughts For Actual Review
##### General
- I might need to go back through and add in notes about general visualisation stuff and bad papers for motivations
- could also help to include a bunch of papers that talk about density visualisation without mentioning uncertainty visualisation in the "context" section. (for now have left those unsorted)

##### Introduction

##### Defining Uncertainty
  - The more taxonomies I read the more convinced I am the "defining uncertainty" is the best one. Not only is it the most complete, but it also identifies WHY some uncertainties are difficult to visualise (specifically which part of the taxonomy) and suggests HOW you can best communicate these uncertainties.
    - A lot of other papers just make a comment like "oh some of these are hard to communicate" but that paper actually goes into detail instead of hand waving the problem away.
    - It also explains why the uncertainty visualisation literature focuses on a specific subset of uncertainty cases.
  - Examples
    - Null Values
    - Variance
    - Hypothesis Testing
    - Distribution
  - Combining uncertainty is also a reccurring theme in these papers
- I think it is also important to highlight if uncertainty is meta information or if it is a variable in of itself. Meta information (such as contiunity, exhangeability, etc) are usually set to specific visual features (e.g. VSUP clearly sees uncertainty as meta information)
- A lot of work that evaluates uncertainty visualisations checks to see if the mean can be gathered from a visualisation, but if we view uncertainty as meta-information, then with a high uncertainty we wouldn't want the mean to be visible
  - Also an example of why the motivation matters for these things. If you don't know "WHY" someone is using a visualisation for uncertainty and just pick the "best" one, you are going to get a visualisation that does a poor job of its task.
  - put uncertainty in the context of other meta-data of variables

##### Visualisation as a form of Uncertainty communication
- A LOT OF the work in uncertainty communication has a sub-field that is visualisation specific
  - Basically I want a brief section highlighting the importance of visualisation and discussing it in relation to other forms of uncertainty communication.
  - especially considering some uncertainty visualisation papers discuss using just a text based method to communicate uncerainty

##### Uncertainty visualisation best practices
  - There is also two kinds of papers, those that compare elements of a plot (like mapping uncertainty to an aesthetic) and those that compare some expression of a distribution (like a scatter plot compared to a box plot).
  - try to map back to david speigleholder paper i.e. uncertainty means lots of different things

# Previous literature review

#####Taxonomy Benefits
- Distinctions at this taxonomy level seem to be important also untangling helps to understand it
  - Failing to communicate the nature of your uncertainty can result in underestimation or overestimation of failure probabilities [@Kiureghian2009]
  - @Gustafson2019 found that the framing of our uncertainty, (i.e. if the source of uncertainty is from a lack of knowledge, approximations, unknown unknowns, or disagreement among parties) was found to not have a detrimental effect on the belief in the estimates, perceived credibility, or behavioural intentions of the decision makers.

##### The benefits of Unvertainty visualisaiton
Visualisations can provide a more complete picture of a risk than numerical summaries alone. Even something as simple as sketching a distribution before recalling statistics or making predictions can greatly increase the accuracy of those measures [@Hullman2018; @Goldstein2014]. While there is some evidence that confidence intervals provided in text form only are less likely to be misinterpreted than graphics [@Savelli2013], text is insufficient to express more complicated aspects of a distribution, such as mass. Expressing uncertainty verbally decreases the perceived reliability and trustworthiness of the source [@VanderBles2020]. Any confusion caused by expressing uncertainty as a visualisation could also be due to a lack of exposure, since @Kay2016 found people repeatedly exposed to the same uncertainty visualisations quickly get better at making judgements. Additionally, visualisation allow for interactive graphics that provide a more in depth understanding of probability [@Potter2009; @Ancker2009] and infographics that make uncertainty more accessible for people with poor numeracy skills [@Ancker2009]. 
**application in energy**
The use of uncertainty in high dimensional environments is especially important in energy data. Large models that incorporate spatial-temporal data from many sources and systems are used to predict energy uses in the short and long term. Understanding how to improve and make better decisions in these models is imperative in both the daily operation of the energy sector as well as in the transition from fossil fuels to clean energy. The energy sector needs better heuristics to make energy supply analysis less costly to conduct [@Stenclik2021], therefore it is an incredibly relevant application of uncertainty visualisation techniques.

##### People dont visualise uncertainty but it is considered important
Despite these benefits, there is evidence that we don't visualise uncertainty as often as we should. A survey conducted by @Hullman2020a found that majority of visualisation authors agreed that expressing uncertainty is important and should be done more often than it currently is, some even agreed that failing to do so is tantamount to fraud. Despite this, only a quarter of respondents included uncertainty in 50% or more of their visualisations [@Hullman2020a]. Meaning participants were convinced that visualising uncertainty is morally important but were able to provide self sufficient reasoning that allows them to avoid doing it. Some economists suggest that visualisation authors are responding to incentives that make it tempting to avoid visualising uncertainty, even if those incentives are based more in perception than reality [@Manski2020]. The study by @Hullman2020a found that the most common reasons authors don't visualise uncertainty despite knowing it's moral importance are: not wanting to overwhelm the audience; an inability to calculate the uncertainty; a lack of access to the uncertainty information; and not wanting to make their data seem questionable [@Hullman2020a].

##### Effect of uncertainty visualisation on decision making
If decision markers are not presented with the uncertainty about an estimate the data analysts have, for all intents and purposes, made the decision for the decision maker. Upon further interviews @Hullman2020a found that authors believed uncertainty would overwhelm the audience and make their data seem questionable because decision makers are unable to understand uncertainty. This belief, while pervasive, is not true. There is some research that suggests laypeople cannot understand complicated concepts in statistical thinking (such as trick questions on hypothesis tests or the difference between Frequentist and Bayesian thinking) [@Hoekstra2014; @Bella2005] but there is a large amount of research suggesting that presenting uncertainty information improves decision making, both experimentally [@Joslyn2012; @Savelli2013; @Kay2016; @Fernandes2018] and in practice [@Al-Kassab2014]. As a matter of fact, doing what many authors currently do (providing only a deterministic outcome with no uncertainty) causes decision makers to be *less* decisive and have completely unbounded expectations on an outcome [@Savelli2013]. This reality cannot be avoided by providing secondary or non-specific information such as explaining calculations [@Joslyn2012], explaining the advantages of a recommendation [@Joslyn2012], or expressing uncertainty in vague terms [@Erev_1990; @Olson_1997], all of which are undesirable for decision makers and lead to measurably worse decisions [@Joslyn2012; @Erev_1990; @Olson_1997]. One of the most popular depictions of uncertainty for decision making is a quantile dotplot, shown in @fig-quantdot. It is important to avoid whittling down the problem **too** much. Providing a categorical decision alone is somewhat useless [@Joslyn2012], and visualising a single estimate is akin to providing a decision or expressing no uncertainty at all.

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-quantdot
#| fig-cap: "This plot depicts an example of a a quantile dotplot that expresses the uncertainty associated with a daily maximum temperature. The probability associated with each temperature is expressed with discrete countable bins and the predicted temperature is expressed with a line. Discretised depictions such as this, make decision making in the face of uncertainty easier for the viewer."
set.seed(1)
dotplot_data <- tibble(temp = 10 + round(rnorm(40, mean=10, sd=2))) 
mid <- round(mean(dotplot_data$temp), 2)
dotplot_data %>%
  ggplot(aes(x=temp)) +
  geom_dotplot(binwidth = 0.75, fill="grey") +
  theme_classic() +
  geom_segment(x=round(mid), xend=round(mid), 
               y=-1, yend=0.7, size=2) +
  geom_label(x=round(mid), y=0.7, label=paste0(mid, "°C"),
             fill = "black", fontface = "bold",
             colour="white") +
  scale_x_continuous(breaks = seq(10, 30, 5), 
                     limits = c(10, 30)) +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank()) +
  labs(x = "Tomorrow's Daily Max Temperature")
```

##### Avoiding uncertainty creates a lack of trust
Not only does communicating uncertainty improve decisions but the mistrust created by communicating certainty in uncertain situations can be exploited. A 6-month survey of anti-mask groups on Facebook during the COVID-19 pandemic showed that the anti-maskers thought carefully about their grammar of graphics and made persuasive visualisations using the same data as pro-mask groups. They did this by exploiting information ignored by the pro-maskers [@Lee2021]. It is understood that deceptive plots can lead viewers to come to incorrect conclusions or significantly overstate effects or risks [@Pandey2015; @Padilla2022] but these incorrect takeaways cannot be mitigated with instructions in how to correctly understand the plot [@Boone2018]. This evidence indicates we are more likely than not to hurt our message when we ignore uncertainty information and trying to raise the general public's plot literacy is an insufficient strategy to curb conspiracy theories and misguided scientific communication.  In direct contrast to this, displaying numerical estimates of uncertainty information has shown to lead to greater trust in predictions [@Joslyn2012; @VanderBles2020]. While @Han2009 found people have more worry when presented with uncertainty regarding health outcomes, this worry is not a bad thing if the concern is warranted given the ambiguous situation.

##### Poeple might simply be chosing to not visualise uncertainty
The disconnect between the research in support of visualising uncertainty and the consensus against it may not be entirely driven by a lack of understanding of the literature. For example, at least one interviewee from the study by @Hullman2020a claimed that expertise implies that the signal being conveyed is significant, but also said they would omit uncertainty if it obfuscated the message they were trying to convey.  Other authors who were capable of calculating and and representing uncertainty well did not do it, and were unable to provide a self-satisfying reason why [@Hullman2020a]. These conflicting motivations are acknowledged in the paper itself where @Hullman2020a says:

> "It is worth noting that many authors seemed confident in stating rationales, as though they perceived them to be truths that do not require examples to demonstrate. It is possible that rationales for omission represent ingrained beliefs more than conclusions authors have drawn from concrete experiences attempting to convey uncertainty".

#### Continuous vs discrete displays
There are other distributional features, such as discreteness, that are important to consider in uncertainty visualisation. There is a reasonable amount of evidence that cumulative displays or discrete displays (such as a quantile dot-plots or histograms) are the best ways to express mass for decision making and probability estimates [@Fernandes2018; @Hofmann2012; Kay2016; @Hullman2018; @kale2019decision].

#### Distributions shoud be in a single plot
Elements from a single distribution should be displayed using a single plot, since displaying the features of one distribution across multiple plots makes the information hard to combine and results in some details (such as the estimate error) being completely ignored [@moritz2017trust; Correll2018].

### Perceptual tasks wrt uncertainty
Not only should we considered the hierarchy in the information we display, but we also need to consider the heuristics that connect some pieces of information to specific elementary tasks. For example @Hofmann2012 showed that polar co-ordinates are more effective than cartesian co-ordinates when considering data that depicts a 360 degree direction (a case where polar co-ordinates has a natural interpretation). Uncertainty also has a natural mapping that should be considered when we express it in a plot. Error is best mapped to fuzziness, location, and colour value; arrangement, size and transparency are an OK second choice; but saturation, hue, orientation and shape are unacceptable and have no intuitive connection to variance [@Maceachren2012]. No only do the graphical elements we map our features to matter, but the direction matters too. Graphical elements that are more fuzzy (fuzziness), further from centre (location), lighter (colour value), poorly arranged (arrangement), smaller (size), more transparent (transparency) are perceived to be more uncertain [@Maceachren2012]. This idea also extends to interval estimation, where questions about probability are best answered with gradients and questions about start and end times are easiest to answer with ambiguation [@Gschwandtnei2016]. Heuristics can work against us just as much as they can work for us. The sine illusion can cause the confidence interval of a smoothed sine curve to seem wider at the peaks than the troughs, causing us to underestimate uncertainty associated with changing values [@Vanderplas2015]. Therefore we should not only keep the hierarchy of information in mind when we map features of our distribution, but also take advantage of these intuitive mappings when we can.